Estimating Sleep Quality Metrics Using Free-Living Accelerometer Data From Thigh-Worn Devices in Comparison to an EEG-Based Sleep Tracking Device

### 2023-07-03 manuscript/manuscript.qmd
bias, loa, ci95% calculated for ml models. Next is biLSTM.

To evaluate our developed models' performance in assessing sleep quality
metrics, we employed Bland-Altman plots and Pearson correlations. These
methods were used to compare the sleep quality summaries calculated by
our models with those obtained from EEG-based ZM sleep quality
summaries. The sleep quality summaries included several metrics:

@@

@@

>>>>

To evaluate our developed models' performance in assessing sleep quality
metrics, we employed Bland-Altman plots and Pearson correlations.
Additionally, we used the Bland-Altman method to assess the agreement
between two measurement techniques. Given the multiple observations per
subject in our dataset, we incorporated a bootstrap procedure to account
for this additional source of variability. We computed the mean
difference (bias) and the limits of agreement as the mean difference
±1.96 times the standard deviation of the differences. The 95%
confidence intervals for the bias and the limits of agreement were
estimated using the bias-corrected and accelerated bootstrap method with
10,000 bootstrap replicates, to ensure our measurements are robust and
account for intra-subject variability. These methods were used to
compare the sleep quality summaries calculated by our models with those
obtained from EEG-based ZM sleep quality summaries. The sleep quality
summaries included the following metrics:

@@

Our models were trained to predict sleep quality metrics using the full
span of accelerometer recordings, beyond the ZM recording duration. This
approach, applied to all models and ZM prediction types, allows for
comprehensive sleep pattern analysis on longitudinal accelerometer data.

@@

# Discussion

The present study embarked on an in-depth exploration of the utility and
comparative performance of different machine learning models in the
analysis of pediatric sleep data. The results offer promising insights
into the role of machine learning in enhancing the precision of sleep
quality metrics.

A significant finding was the impact of median filters on the sleep
metrics. Implementing 5-minute and 10-minute median filters led to
notable changes in key sleep parameters. In particular, the Total Sleep
Time (TST), Sleep Efficiency (SE), and Latency to Persistent Sleep (LPS)
increased, suggesting that the filters may categorize some instances of
wakefulness as sleep and smooth out brief awakenings. This underscores
the potential of median filters in refining sleep data and mitigating
the potential misclassification of sleep states.

Interestingly, the most marked change was observed in Wake After Sleep
Onset (WASO), with a substantial reduction in the filtered data. This
reduction, coupled with the decrease in the number of awakenings,
implies a significant role of median filters in sleep-wake cycle
characterization, potentially enhancing the reliability of sleep
assessment.

The machine learning models demonstrated high overall performance on an
epoch-to-epoch basis, with the XGBoost model marginally outperforming
the others. However, the models had more varied performance in
classifying sleep conditions, as evidenced by the sleep performance
metrics. The relatively lower Specificity values across all models
suggest a common challenge in correctly classifying negative instances.

In the precision-recall and ROC curve analyses, the Decision Tree model
showed superior precision-recall AUC, indicating strong predictive
accuracy. In contrast, the XGBoost model consistently demonstrated a
high ROC AUC, indicating a robust ability to differentiate between
classes. However, it's worth noting that the Neural Network model
generally showed weaker performance in these areas.

The three-class biLSTM multiclassifier showed overall good performance,
further supporting the potential of machine learning in sleep study.
However, the absence of Specificity values limits a complete assessment
of its performance, calling for further exploration in future studies.

In conclusion, this study provides compelling evidence of the utility of
machine learning models and median filters in enhancing the precision
and reliability of pediatric sleep assessment. However, more research is
needed to overcome the challenge of correctly classifying negative
instances and to fully understand the performance of these models across
all classes and conditions. The findings underscore the potential of
machine learning in advancing sleep medicine, paving the way for more
accurate, reliable, and personalized sleep assessment in pediatric
populations.

### 2023-07-03 manuscript/manuscript.tex
bias, loa, ci95% calculated for ml models. Next is biLSTM.

\begin{keyword}
    Sleep \sep Accelerometry \sep EEG \sep Machine learning \sep 
    Sleep quality metrics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, frame hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

To evaluate our developed models' performance in assessing sleep quality
metrics, we employed Bland-Altman plots and Pearson correlations. These
methods were used to compare the sleep quality summaries calculated by
our models with those obtained from EEG-based ZM sleep quality
summaries. The sleep quality summaries included several metrics:

@@

>>>>

\begin{keyword}
    Sleep \sep Accelerometry \sep EEG \sep Machine learning \sep 
    Sleep quality metrics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced]}{\end{tcolorbox}}\fi

@@

To evaluate our developed models' performance in assessing sleep quality
metrics, we employed Bland-Altman plots and Pearson correlations.
Additionally, we used the Bland-Altman method to assess the agreement
between two measurement techniques. Given the multiple observations per
subject in our dataset, we incorporated a bootstrap procedure to account
for this additional source of variability. We computed the mean
difference (bias) and the limits of agreement as the mean difference
±1.96 times the standard deviation of the differences. The 95\%
confidence intervals for the bias and the limits of agreement were
estimated using the bias-corrected and accelerated bootstrap method with
10,000 bootstrap replicates, to ensure our measurements are robust and
account for intra-subject variability. These methods were used to
compare the sleep quality summaries calculated by our models with those
obtained from EEG-based ZM sleep quality summaries. The sleep quality
summaries included the following metrics:

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The present study embarked on an in-depth exploration of the utility and
comparative performance of different machine learning models in the
analysis of pediatric sleep data. The results offer promising insights
into the role of machine learning in enhancing the precision of sleep
quality metrics.

A significant finding was the impact of median filters on the sleep
metrics. Implementing 5-minute and 10-minute median filters led to
notable changes in key sleep parameters. In particular, the Total Sleep
Time (TST), Sleep Efficiency (SE), and Latency to Persistent Sleep (LPS)
increased, suggesting that the filters may categorize some instances of
wakefulness as sleep and smooth out brief awakenings. This underscores
the potential of median filters in refining sleep data and mitigating
the potential misclassification of sleep states.

Interestingly, the most marked change was observed in Wake After Sleep
Onset (WASO), with a substantial reduction in the filtered data. This
reduction, coupled with the decrease in the number of awakenings,
implies a significant role of median filters in sleep-wake cycle
characterization, potentially enhancing the reliability of sleep
assessment.

The machine learning models demonstrated high overall performance on an
epoch-to-epoch basis, with the XGBoost model marginally outperforming
the others. However, the models had more varied performance in
classifying sleep conditions, as evidenced by the sleep performance
metrics. The relatively lower Specificity values across all models
suggest a common challenge in correctly classifying negative instances.

In the precision-recall and ROC curve analyses, the Decision Tree model
showed superior precision-recall AUC, indicating strong predictive
accuracy. In contrast, the XGBoost model consistently demonstrated a
high ROC AUC, indicating a robust ability to differentiate between
classes. However, it's worth noting that the Neural Network model
generally showed weaker performance in these areas.

The three-class biLSTM multiclassifier showed overall good performance,
further supporting the potential of machine learning in sleep study.
However, the absence of Specificity values limits a complete assessment
of its performance, calling for further exploration in future studies.

In conclusion, this study provides compelling evidence of the utility of
machine learning models and median filters in enhancing the precision
and reliability of pediatric sleep assessment. However, more research is
needed to overcome the challenge of correctly classifying negative
instances and to fully understand the performance of these models across
all classes and conditions. The findings underscore the potential of
machine learning in advancing sleep medicine, paving the way for more
accurate, reliable, and personalized sleep assessment in pediatric
populations.
